---
title: "Assignment 1 Report"
author: "Winson Huang"
date: "January 07, 2025"
output: html_document
---
[GitHub Repository](https://github.com/winsonnhuang/BIOL432_Assignment1)

# Introduction
This report analyzes hypothetical limb measurements and volumes using R scripts to summarize statistics using tables and figures.

```{r setup, include=FALSE}
# Load the required libraries for data manipulation, visualization and table formatting
library(dplyr)
library(ggplot2)
library(knitr)
```

```{r run-scripts, echo=F}
# Executing the R scripts to generate hypothetical data and caluclating limb volumes
source("dataGenerato.R")
source("volumeEstimato.R")
```

```{r load-data, echo=F}
# Reading the processed data from the CSV file
data <- read.csv("measurements.csv")
```

```{r sort-data, echo=F}
# Sorting the Data by species name, observer name, and volume
sorted_data <- data %>%
  arrange(Organism, Observer, Volume)
```

# Average Species Volume

```{r average-volume, echo=F}
# Group data by species and calculate the mean volume for each species
avg_volume <- data %>%
  group_by(Organism) %>%
  summarize(Average_Volume = mean(Volume, na.rm= TRUE))
# Displays the average volume table
kable(avg_volume)
```


# Number of Observations for each Combination of Species and Observer

```{r count-observations, echo=F}
# Group data by species and observer and count the number of rows in each group
obs_count <- data %>%
  group_by(Organism, Observer) %>%
  summarize(Count = n(), .groups = "drop")
# Displays the observation count table
kable(obs_count, caption = "Observations by Species and Observer")
```

# Boxplot to compare key features of the distributions of volumes for each species

```{r boxplot-volume, echo=F}
# Creating a boxplot to visualize the distributions
ggplot(data, aes(x = Organism, y = Volume)) +
  geom_boxplot() +
  labs(x = "Species", y = "Volume")
```

# Histograms to compare distribution of limb volume for each species

```{r histogram, echo=F}
# Creating histograms to show the frequency distribution
ggplot(data, aes(x = Volume)) +
  geom_histogram(binwidth = 10, color = "skyblue") +
  facet_wrap(~ Organism, scales = "fixed") +
  labs(x = "Volume", y = "Frequency")
```